Make_mash_U_hyp

### 2020-07-09 R/Make_mash_U_hyp.Rmd
Try rescaling all FL50 gulf effects to fall on same after-solstice scale

U_hyp must be matrices. They must be symmetric, in the same order as Bhat etc, and with the same # rows and cols as Bhat etc.

@@

@@

m <- mash_standard_run(path = path, numSNPs = numSNPs, U_ed = U_ed,
                       U_hyp = U_h1, suffix = suffix, saveoutput = TRUE)
```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
m <- readRDS(file.path(path, "Strong_Effects30000SNPs_gulf_uhyp1.rds"))
suffix = "gulf_uhyp1"
mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
mash_plot_Ulist(m, range = get_U_by_mass(m), saveoutput = TRUE, 
                  suffix = suffix)
mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 8)

@@

>>>>

U_hyp must be numeric matrices. They must be symmetric, square, in the same order as Bhat etc, and with the same # rows and cols as Bhat etc.

They can't contain NaN, NA, Inf values. And it might help if they're positive definite.

That's all the info in the mash package that I can find.

I suppose that I want them to be on the same scale as the mash input data, not the same scale as whatever weather or distance data they came from. This is a sense in which correlation matrices would be better. Take a look at the U_ed. Rescale to a similar scale as ED_PCA_1. But then...

@@

U_h2 <- list(Udist = Ugulfhyp$U_dist, UGDD = Ugulfhyp$U_CGDD_12C)

@@

m <- mash_standard_run(path = path, numSNPs = numSNPs, U_ed = U_ed,
                       U_hyp = U_h1, suffix = suffix, saveoutput = TRUE)
m <- mash_standard_run(path = path, numSNPs = numSNPs, U_ed = U_ed,
                       U_hyp = U_h2, suffix = suffix, saveoutput = TRUE)
```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
m <- readRDS(file.path(path, "Strong_Effects30000SNPs_gulf_uhyp1.rds"))
suffix = "gulf_uhyp1"
mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
mash_plot_Ulist(m, range = get_U_by_mass(m))
mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 8)

@@

```{r}
load_all("~/Github/switchgrassGWAS")
fitted_g <- list(Ulist = u_ed)
m_mock <- list(fitted_g = fitted_g)
mash_plot_Ulist(m_mock, limits = FALSE)

bhat %>% as_tibble() %>%
  colMeans()
summary(bhat)

# CLMB_FL50 KBSM_FL50 KING_FL50 PKLE_FL50 TMPL_FL50 
240.7987 - 173  
261.9135 - 173   
209.8265 - 173  
210.2195 - 173  
240.0100 - 173

B_hat_strong_df_30000topSNPs_gulf_solstice <- bhat %>%
  mutate(Bhat_CLMB_FL50 = Bhat_CLMB_FL50 + (240.7987 - 173),
         Bhat_KING_FL50 = Bhat_KING_FL50 + (209.8265 - 173 ),
         Bhat_KBSM_FL50 = Bhat_KBSM_FL50 + (261.9135 - 173),
         Bhat_PKLE_FL50 = Bhat_PKLE_FL50 + (210.2195 - 173),
         Bhat_TMPL_FL50 = Bhat_TMPL_FL50 + (240.0100 - 173),
         ) #%>%
  #summary()
write_rds(B_hat_strong_df_30000topSNPs_gulf_solstice, file.path(path, "B_hat_strong_df_30000topSNPs_gulf_solstice.rds"))

B_hat_random_df_30000topSNPs_gulf_solstice <- B_hat_random_df_30000topSNPs_gulf_uhyp1 %>%
  mutate(Bhat_CLMB_FL50 = Bhat_CLMB_FL50 + (240.7987 - 173),
         Bhat_KING_FL50 = Bhat_KING_FL50 + (209.8265 - 173 ),
         Bhat_KBSM_FL50 = Bhat_KBSM_FL50 + (261.9135 - 173),
         Bhat_PKLE_FL50 = Bhat_PKLE_FL50 + (210.2195 - 173),
         Bhat_TMPL_FL50 = Bhat_TMPL_FL50 + (240.0100 - 173),
         ) 
write_rds(B_hat_random_df_30000topSNPs_gulf_solstice, file.path(path, "B_hat_random_df_30000topSNPs_gulf_solstice.rds"))
```
what is each site's mean FL50? Just for Gulf plants.

For interpretability of the mash results... may want to scale all these effects so that 0 is, say, the summer solstice? Instead of the mean flowering time at that site for that pop.
```{r}
phe_gulf <- phenotypes$Gulf %>%
  select(PLANT_ID, ends_with("FL50"))

sd(phe_gulf$CLMB_FL50, na.rm = TRUE)
sd(phe_gulf$KBSM_FL50, na.rm = TRUE)
sd(phe_gulf$KING_FL50, na.rm = TRUE)
sd(phe_gulf$PKLE_FL50, na.rm = TRUE)
sd(phe_gulf$TMPL_FL50, na.rm = TRUE)

summary(phe_gulf$CLMB_FL50, na.rm = TRUE)
summary(phe_gulf$KBSM_FL50, na.rm = TRUE)
summary(phe_gulf$KING_FL50, na.rm = TRUE)
summary(phe_gulf$PKLE_FL50, na.rm = TRUE)
summary(phe_gulf$TMPL_FL50, na.rm = TRUE)
summary(phe_gulf)

```
CLMB_FL50 KBSM_FL50 KING_FL50 PKLE_FL50 TMPL_FL50 
240.7987   261.9135  209.8265  210.2195 240.0100

241 +/- 18.6
262 +/- 10.5
210 +/- 34
210 +/- 35
240 +/- 28
